Methods for Large Scale Learning to Rank

### 2014-03-05 report/Chapters/Background.tex
uitleg NP-Hard probleem pairwise ranking

\chapter{Approaches to Learning-to-Rank}
\section{Pointwise Approach}
The pointwise approach can be seen as the most straightforward way of using machine learning for ranking. Pointwise Learning-to-Rank methods directly apply machine learning methods to the ranking problem by observing each document in isolation. They can be subdivided in the following approaches:
	\begin{enumerate}
	\item regression-based, which estimate the relevance of a considered document using a regression model.
	\item classification-based, which classify the relevance category of the document using a classification model.
	\item ordinal regression-based, which classify the relevance category of the document using a classification model in such a way that the order of relevance categories is taken into account. 
	\end{enumerate}
\section{Pairwise Approach}
Pointwise Learning-to-Rank methods have the drawback that they optimise real valued expected relevance, while evaluation metrics like \ac{nDCG} and \ac{ERR} are only impacted by a change in expected relevance when that change impacts a pairwise preference. The pairwise approach solves this drawback of the pointwise approach by regarding ranking as pairwise classification. Aggregating a set of predicted pairwise preferences into the corresponding optimal rank is shown to be a NP-Hard problem. 
\section{Listwise Approach}
Listwise ranking optimises the actual evaluation metric. The learner learns to predict an actual ranking itself without using an intermediate step like in pointwise or pairwise Learning-to-Rank. The main challenge in this approach is that most evaluation metrics are not differentiable. \ac{MAP}, \ac{ERR} and \ac{nDCG} are non-differentiable, non-convex and discontinuous functions, what makes them very hard to optimize.\\

>>>>

\chapter{Approaches to Learning-to-Rank}
\section{Pointwise Approach}
The pointwise approach can be seen as the most straightforward way of using machine learning for ranking. Pointwise Learning-to-Rank methods directly apply machine learning methods to the ranking problem by observing each document in isolation. They can be subdivided in the following approaches:
	\begin{enumerate}
	\item regression-based, which estimate the relevance of a considered document using a regression model.
	\item classification-based, which classify the relevance category of the document using a classification model.
	\item ordinal regression-based, which classify the relevance category of the document using a classification model in such a way that the order of relevance categories is taken into account. 
	\end{enumerate}
\section{Pairwise Approach}
Pointwise Learning-to-Rank methods have the drawback that they optimise real valued expected relevance, while evaluation metrics like \ac{nDCG} and \ac{ERR} are only impacted by a change in expected relevance when that change impacts a pairwise preference. The pairwise approach solves this drawback of the pointwise approach by regarding ranking as pairwise classification.\\

Aggregating a set of predicted pairwise preferences into the corresponding optimal rank is shown to be a NP-Hard problem \cite{Feldman2012}. An often used solution to this problem is to upper bound the number of classification mistakes by an easy to optimise function \cite{Bartlett2006}.
\section{Listwise Approach}
Listwise ranking optimises the actual evaluation metric. The learner learns to predict an actual ranking itself without using an intermediate step like in pointwise or pairwise Learning-to-Rank. The main challenge in this approach is that most evaluation metrics are not differentiable. \ac{MAP}, \ac{ERR} and \ac{nDCG} are non-differentiable, non-convex and discontinuous functions, what makes them very hard to optimize.\\

### 2014-03-05 report/RelatedWork.tex
uitleg NP-Hard probleem pairwise ranking

Tyree et al. \cite{Tyree2011} describe a way of parallelising \ac{GBDT} models for Learning-to-Rank where the construction of the regression trees themselves are parallelised, based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}. Decision trees are build layer-by-layer where the calculations needed for building each new layer in the tree are divided over the nodes using a master-worker paradigm. The data is partitioned of the workers, who compress their share into histograms and send these to the master. The master uses those histograms to approximate the split and build the next layer. The master than communicates this new layer to the workers who can use this new layer to compute new histograms. this process is repeated until the tree depth limit is reached. The tree construction algorithm parallelised with described master-worker method is \ac{CART} \cite{Breiman1984}. Speed-up experiments on the LETOR and the Yahoo! Learning to Rank challenge data sets were performed. This parallel \ac{CART}-tree building approach showed speed-up of up to 42x on shared memory machines and up to 25x on distributed memory machines.
\section{Deep Stacking Networks}
A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable. \ac{DSN} is based on the stacked generalization ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack. Each layer in the stack learns has the task of learning the weights of the inputs of that layer. The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.
\section{Alternating Direction Method of Multipliers}
Duh et al. \cite{Duh2011} propose the use of \ac{ADMM} for the Learning-to-Rank task. \ac{ADMM} is a general optimization method that solves problems of the form
\begin{alignat*}{2}
\text{minimize }   &  f(x) + g(x) \\
\text{subject to } &  Ax + Bz = c
\end{alignat*}
by updating $x$ and $z$ in an alternating fashion. It holds the nice properties that it can be executed in parallel and that it allows for incremental model updates without full retraining. Duh et al. \cite{Duh2011} show how to use \ac{ADMM} to train a RankSVM model in parallel. Experiments showed the \ac{ADMM} implementation to achieve a 13.1x training time speed-up on 72 workers, compared to training on a single worker.\\

>>>>

Tyree et al. \cite{Tyree2011} describe a way of parallelising \ac{GBDT} models for Learning-to-Rank where the boosting step is still executed sequentially, instead the construction of the regression trees themselves are parallelised. The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}. Decision trees are build layer-by-layer where the calculations needed for building each new layer in the tree are divided over the nodes using a master-worker paradigm. The data is partitioned of the workers, who compress their share into histograms and send these to the master. The master uses those histograms to approximate the split and build the next layer. The master than communicates this new layer to the workers who can use this new layer to compute new histograms. this process is repeated until the tree depth limit is reached. The tree construction algorithm parallelised with described master-worker method is \ac{CART} \cite{Breiman1984}. Speed-up experiments on the LETOR and the Yahoo! Learning to Rank challenge data sets were performed. This parallel \ac{CART}-tree building approach showed speed-up of up to 42x on shared memory machines and up to 25x on distributed memory machines.
\section{Deep Stacking Networks}
A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable. \ac{DSN} is based on the stacked generalization ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack. Each layer in the stack learns has the task of learning the weights of the inputs of that layer. The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.
\section{Alternating Direction Method of Multipliers}
Duh et al. \cite{Duh2011} propose the use of \ac{ADMM} for the Learning-to-Rank task. \ac{ADMM} is a general optimization method that solves problems of the form
\begin{alignat*}{2}
\text{minimize }   &  f(x) + g(x) \\
\text{subject to } &  Ax + Bz = c
\end{alignat*}
by updating $x$ and $z$ in an alternating fashion. It holds the nice properties that it can be executed in parallel and that it allows for incremental model updates without full retraining. Duh et al. \cite{Duh2011} show how to use \ac{ADMM} to train a RankSVM model in parallel. Experiments showed the \ac{ADMM} implementation to achieve a 13.1x training time speed-up on 72 workers, compared to training on a single worker.\\
